---
title: "P8105 HW1"
author: "Zakari L. Billo"
date: "09/20/2025"
output: html_document
---

Load libararies

```{r, echo = FALSE}
library(tidyverse)
library(moderndive)
```

# Problem 1

```{r import step and basic anal}
data("early_january_weather") #load dataset

weather_df <- early_january_weather |>
  janitor::clean_names() #clean names in new df, preserve original
  
summary(weather_df) #view summary
print(nrow(weather_df)) #358 rows
print(ncol(weather_df)) #15 col

print(mean(pull(weather_df, temp))) #39.58212
```

The dataset "early_january_weather" (renamed weather_df) contains 15 variables (reflecting 15 columns) and 358 observations (reflecting 358 rows):
* `origin`: airport code (EWR)
* `year`: 2013
* `month`: january
* `day`: january 1-15th
* `hour`: 12:00 am - 23:00 pm
* `temp`: mean = 39.58212 (ªF)
* `dewp`: mean = 28.06 (ªF)
* `humid`: mean = 65.48 (%)
* `wind_dir`: mean = 208.2 (degrees from north)
* `wind_speed`: mean = 8.226 (mph)
* `wind_gust`: mean = 22.53 (mph)
* `precip`: mean = 0.002039 (inches)
* `pressure`: mean = 1023 (millibars)
* `visib`: mean = 8.515 (miles)
* `time_hour`: range jan 1st at 01:00:00 am to jan 15th at 23:00:00 pm

As requested, the mean temperature was 39.58212 (ªF).

```{r}
# scatterplot of `temp` vs. `time` w `humid` as color
ggplot(weather_df, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs( #labels
    title = "Temperature (ªF) vs Time (Hour)",
    x = "Hour",
    y = "Temperature (°F)",
    color = "Humidity (%)"
  ) 

#export plot
ggsave('temp_hour_scatter.jpeg')
```

The scatterplot of `temp` (ªF) vs. `time_hour` (hours) shows that the temperature rises during the days, peaking in the afternoon and lowering by the evening. The minimum temperatures throughout the days steadily increased throughout the month of January. The relative humidity tends was low during this time period, but increased significantly around Jan 14 when temperatures rose. 
